Fast Einsum Implementation Using Parallelization, Vectorization, and More\\\large Algorithm Engineering 2025 Project Paper

### 2025-03-04 answers/chapters/01/01.tex
restructuring

~\\
\includegraphics[width=\textwidth]{page_15.png}

~\\
Let us proof the statement
\[
    \int_{0}^{1}\frac{1}{1+x^2} dx = \frac{\pi}{4}
\]

@@

Even better is the approximation by parabolas, which pass through the end points of the interval as well as the mid point. This is called the \em Simpson Rule\em.
It reads as follows:
\[
    \int_{t}^{t+h} f(x) dx \ \approx \ \frac{h}{6} \Big[ f(t) + 4f(t+\frac{h}{2}) + f(t+h) \Big] \ \ \ \ \ .
\]

@@

\[
    \int_{a}^{b} f(x) dx \ \ \approx \sum_{(t, t+h) \in P} I_{(t, t+h)}(f) \ \ \ \ \ ,
\]

where \(I_{(t, t+h)}(f)\) tries to approximate \(\int_{t}^{t+h} f(x) dx\).

@@

\subsubsection{Memory Hierarchy}
The inverse correlation between speed and volume of different memory types war presented in the book.
This lead to the memory hierarchy, where we try exploit the benefits of both worlds be having big, slow memory, and a smaller amount of fast memory, which we the computer tries to employ as much as possible.

>>>>

\begin{figure}[h]
    \includegraphics[width=\textwidth]{page_15.png}
    \caption{Slide 15 from lecture 1}
\end{figure}

~\\
Let us prove the statement
\[
    \int_{0}^{1}\frac{1}{1+x^2} dx = \frac{\pi}{4}
\]

@@

\noindent
Even better is the approximation by parabolas, which pass through the end points of the interval as well as the mid point. This is called the \em Simpson Rule\em.
It reads as follows:
\[
    \int_{t}^{t+h} f(x) dx \ \approx \ \frac{h}{6} \Big[ f(t) + 4f(t+\frac{h}{2}) + f(t+h) \Big] \ \ \ \ \ .
\]

@@

\[
    \int_{a}^{b} f(x) dx \ \ \approx \sum_{(t, t+h) \in P} I_{(t, t+h)}(f) \ \ \ \ \ ,
\]
\noindent
where \(I_{(t, t+h)}(f)\) tries to approximate \(\int_{t}^{t+h} f(x) dx\).

@@

\subsubsection{Memory Hierarchy}
The inverse correlation between speed and volume of different memory types was presented in the book.
This lead to the memory hierarchy, where we try exploit the benefits of both worlds be having big, slow memory, and a smaller amount of fast memory, which we the computer tries to employ as much as possible.

### 2025-03-04 answers/chapters/11/11.tex
restructuring

~\\
First, we should keep in mind the page size of the SSD, which typically is 4KB. Hence, small read and write operation below this mark will have the same cost of a 4KB one, and in case of a write operation we may not fully utilized the space as the page we are writing to flash itself is not fully utilized. Thus, we should combine smaller IO operations in two fewer bigger ones, provided that these operations are sequential, which directly leads us to the second conclusion:

@@

~\\
\subsection{Designing SSD-friendly Applications for Better Application Performance and Higher IO Efficiency}
This paper introduces the reader to the core functioning of SSDs and presents different optimization techniques to improve performance or longevity of the SSD. One optimization presented in the paper directly follows on from the previous discussion. The other one I chose to present analyzes the page distribution inside the SSD and its effect on performance.

\begin{description}
    \item[Seperate hot data from cold data] We have already established that we should group related data. But additionally, or rather more importantly, we should group hot data sequentially (i.e. data accessed very often) as well as cold data. The optimum for all data would be a smooth transition between these two extrema. The reason for this is very similar as before, as when accessing data from the SSD we also access unwanted data due the teh page size. Optimally, we also require this additional data, otherwise this data section is accessed redundantly. To maximize the chance of the first scenario, we organize our data in the proposed manner.
    \item[Avoid full SSD usage] When we assume that the used pages distributed equally among the entire SSD, and the SSD is occupied by a factor of $A$ ($A = 0.5 \implies$ half the pages are in use), then for each block on average we have $\frac{no\_occupied\_pages\_in\_block}{no\_pages\_in\_block} = A$. Thus, we see that the garbage collector will have to do way less data movement when $A$ is low. In fact, $\frac{1}{1-A}$ blocks will need to be considered when trying to free a block on average, and also note that for each of these blocks, $A \cdot no\_pages\_in\_block$ pages will need to be copied. Thus, we face a double penalty for bigger values of $A$ ($\frac{A \cdot no\_pages\_in\_block}{1-A}$ will need to be moved; in the paper $P$ is used for $no\_pages\_in\_block$).
\end{description}
\end{document}

>>>>

~\\
First, we should keep in mind the page size of the SSD, which typically is 4KB. Hence, small read and write operation below this mark will have the same cost of a 4KB one, and in case of a write operation we may not fully utilize the space as the page we are writing to flash itself is not fully utilized. Thus, we should combine smaller IO operations in two fewer bigger ones, provided that these operations are sequential, which directly leads us to the second conclusion:

@@

\bigskip
\subsection{Designing SSD-friendly Applications for Better Application Performance and Higher IO Efficiency}
\href{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=7552037}{This paper} introduces the reader to the core functioning of SSDs and presents different optimization techniques to improve performance or longevity of the SSD. One optimization presented in the paper directly follows on from the previous discussion. The other one I chose to present analyzes the page distribution inside the SSD and its effect on performance.

\begin{description}
    \item[Seperate hot data from cold data] We have already established that we should group related data. But additionally, or rather more importantly, we should group hot data sequentially (i.e. data accessed very often) as well as cold data. The optimum for all data would be a smooth transition between these two extrema. The reason for this is very similar as before, as when accessing data from the SSD we also access unwanted data due the page size. Optimally, we also require this additional data, otherwise this data section is accessed redundantly. To maximize the chance of the first scenario, we organize our data in the proposed manner.
    \item[Avoid full SSD usage] When we assume that the used pages distributed equally among the entire SSD, and the SSD is occupied by a factor of $A$ ($A = 0.5 \implies$ half the pages are in use), then for each block on average we have $\frac{no\_occupied\_pages\_in\_block}{no\_pages\_in\_block} = A$. Thus, we see that the garbage collector will have to do way less data movement when $A$ is low. In fact, $\frac{1}{1-A}$ blocks will need to be considered when trying to free a block on average, and also note that for each of these blocks, $A \cdot no\_pages\_in\_block$ pages will need to be copied. Thus, we face a double penalty for bigger values of $A$ ($\frac{A \cdot no\_pages\_in\_block}{1-A}$ will need to be moved; in the paper $P$ is used for $no\_pages\_in\_block$).
\end{description}
\end{document}

### 2025-03-04 paper/paper.tex
restructuring

            #pragma omp simd
            for (size_t k = 0; k < inner_dim; ++k) {
                sum += lhs_row[k] * rhs_col[k];
            }
            result_row[j] = sum;
        }
    }
}
\end{lstlisting}
\\\\
We face the \textbf{exact} problems we faced at the reduction method, and my responses are the same, we just keep the implementation simple for the sake of demonstration. This function could be manually vectorized as well, but the compiler should do a decent job at optimizing this function, and hence we shouldn't face a big performance penalty.

>>>>

            #pragma omp simd
            for (size_t k = 0; k < inner_dim; ++k) {
                sum += lhs_row[k] * rhs_col[k];
            }
            result_row[j] = sum;
        }
    }
}
\end{lstlisting}
\\\\
We face the \textbf{exact} problems we faced at the reduction method, and my responses are the same, we just keep the implementation simple for the sake of demonstration. This function can be manually vectorized as well, however the compiler should do a decent job at optimizing this function, and hence we shouldn't face a big performance penalty. For \texttt{float} and \texttt{double} vectorized code can be found in specialized template methods below.
